---
title: "rsibal_OriginalHomeworkCode_01"
author: "Ritika"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: true
      number_sections: true
      toc_float: true
      theme: cerulean
---

*I really liked how you changed the theme of your html doc. Will definitely be implementing it in my own code.*

# Summary

This is Ritika Sibal's R Markdown File for Modules 3,4 & 5. Both work and notes have been separated by modules where relevant. Challenges location in second half of file

## Module 4 Notes

A matrix is a collection of vectors stored together, can only have one data type

```{r Module 4}
m<-matrix(data=c(1,2,3,4), nrow=2, ncol=2)
```

> NOTE Matrices are filled column wise. If you want them to be filled row wise use the *byrow=TRUE* flag (last input argument)

You can also bind vectors of same length togther row wise *rbind(vector1, vector2)* or col wise *cbind(vector1, vector2)*

### Module 4 Notes 1

```{r Module 4 Notes 1}
v1 <- c(1, 2, 3, 4)
v2 <- c(6, 7, 8, 9)
m<-cbind(v1, v2)
```

Use the following functions for Metadata in matrices: *class()*, *dim()*, *names()*, *rownames()*, *colnames()* with *dim()* returns a 1x2 vector with number of rows in index 1 and number of cols in index 2

*str()* is used to compactly display the internal structure of an R object. Example seen below:

### Module 4 Notes 2

```{r Module 4 Notes 2}
str(m)
```

Arrays, on the other hand comprise of both vectors (1D) and matrices (2D). Must have same atomic data class. An example of a 3D array is seen below

### Module 4 Notes 3

```{r Module 4 Notes 3}
a <- array(data = 1:90, dim = c(5, 6, 3))
a
```

To index into an array, vector or matrix you can use square brackets *[]* with the index/indices within the brackets (can highlight a subset too)

### Module 4 Notes 4

```{r Module 4 Notes 4}
v <- 1:100 
v[1:15] # First 15 elements
v[c(2, 4, 6, 8, 10)] # Set of values to subset, where the values in the brackets reflect the indices to subset, not the actual values
```

For matrices you need to double index (row and then col) *m[row #, col#]*

You can also assign values in the matrix. Remember the format is *matrix[row,col] \<- value*

The following example was interesting to me, where the first row from columns 4:8 was assigned the first vector value and the second row was assigned the second.

Also note that the numeric vector becomes of type char when setting any row/col value to a char/string

### Module 4 Notes 5

```{r Module 4 Notes 5}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
m[2:5, 4:8] <- c(20, 19, 18, 17)
m[2:5, 4:8] <- matrix(data = c(20:1), nrow = 4, ncol = 5, byrow = TRUE)
m[, 8] <- c("a", "b")
```

Lists, can be used to group together a mix of R structures (list can contain a matrix, string array etc). Extracting requires double brackets e.g *l[[2]]* to index the matrix, and *l[[2]][2, 6]* to index into the matrix.

You can also name elements in the list using the *names()* function. EX: *names(list) \<- c("name 1", "name 2", "name 3")* and use *list\$name1* to slice that value in the list

### Module 4 Notes 6

```{r Module 4 Notes 6}
s <- c("this", "is", "a", "vector", "of", "strings")  # here we have a vector
m <- matrix(data = 1:40, nrow = 5, ncol = 8)  # this is a matrix
b <- FALSE  #and this is a logical statement
l <- list(s, m, b)  # this puts them all together as a list
l[1:2]  # here we're slicing our list so it just has the first and second elements. LIST SLICING
```

Data frames are good at showing tabular data. Use *cbind()* or *rbind()* to add more rows or cols.

> NOTE that the frame is more efficient for strings because internally R assigns int values to the unique string (like an enum)

R had special values, keep track of them!

NA (for not available) is used for missing data. Many statistical functions offer the possibility to include as an argument na.rm=TRUE ("remove NAs") so that NAs are excluded from a calculation. Inf (and -Inf) is used when the result of a numerical calculation is too extreme for R to express NaN (for not a number) is used when R cannot express the results of a calculation , e.g., when you try to take the square root of a negative number

## Module 4 Challenge

### Module 4 Challenge 1

*I liked how you commented out what you were doing with every line. Made it easy to interpret and follow along.*

```{r Module 4 Challenge 1}
u_string="Stately, plump Buck Mulligan came from the stairhead, bearing a bowl of lather on which a mirror and a razor lay crossed."
# gsub removes the punctuation!
u_string<-gsub("[[:punct:]]","",u_string)
# I choose to use strsplit() instead of (the suggested) str_split() based on the information found
# https://stackoverflow.com/questions/7367284/inconsistent-behavior-between-str-split-and-strsplit

# Separate the string using a space deliminator
u_vector<-strsplit(u_string, " ")[[1]]
# Alternatively, I could have done:
# u_string<-gsub("[[:punct:]]","",u_string) %>%  strsplit(, " ")[[1]]
u_vector[seq(from=3,to=length(u_vector),by=3)]
```

*I like how you used the pipe connectors. Really organized, really simple.*

### Module 4 Challenge 2

```{r Module 4 Challenge 2}
m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE)
x<-m[,c(2,3,6)] # Extract columns 2,3 & 6
x<-m[6:8,] # Extract rows 6 to 8 
x<-m[2:6, 2:9] # Extract row2col2 to row6col9
```

*I don't know if this is just user error on my part, but when I run the code, it doesn't print out the results.*

### Module 4 Challenge 3

```{r Module 4 Challenge 3}
# Create the array
a<-array(data=400:1, dim=c(5,5,4,4))
# Get all the indexed values
a[1,1,1,2]
a[2,3,2,]
a[1:5,1:5,3,3]
```

### Module 4 Challenge 4

```{r Module 4 Challenge 4}
# This section was just copying the example and changing the names of the primates.
Platyrrhini <- c("Cebidae", "Atelidae", "Pitheciidae")
Catarrhini <- c("Cercopithecidae", "Hylobatidae", "Hominidae")
Anthropoidea <- list(Platyrrhini, Catarrhini)
names(Anthropoidea) <- list("Platyrrhini", "Catarrhini")

Tarsioidea <- c("Tarsiidae")
names(Tarsioidea) <- list("Tarsiidae")

Haplorhini <- list(Anthropoidea, Tarsioidea)
names(Haplorhini) <- list("Anthropoidea", "Tarsioidea")

Lemuroidea <- c("Cheirogaleidae", "Lepilemuridae", "Indriidae", "Lemuridae",
    "Daubentoniidae")
Lorisoidea <- c("Lorisidae", "Galagidae")
Strepsirhini <- list(Lemuroidea, Lorisoidea)
names(Strepsirhini) <- list("Lemuroidea", "Lorisoidea")

Primates <- list(Haplorhini, Strepsirhini)
names(Primates) <- list("Haplorhini", "Strepsirhini")
Primates
```

*I really liked how you did this challenge. It definitely looks more organized than mine. I was having trouble distinguishing the lists so I'll be sure to implement some of these techniques into my work.*

### Module 4 Challenge 5

```{r Module 4 Challenge 5}
# use byrow=TRUE to make sure we are loading by ROW
m3 <- matrix(c(3, 0, 1 ,23, 1, 2, 33, 1, 1, 42, 0, 1, 41, 0, 2), 5, 3, byrow=TRUE)
# Coerce the matrix to a data frame.
df_m3<-as.data.frame(m3)
# Coerce the second column to be logical-valued. The follow is evaluates df_m3[,2]==1 first and sets it to df_m3[,2]
df_m3[,2]=df_m3[,2]==1
# Coerce the third column to be factor-valued
df_m3[,3]=as.factor(df_m3[,3])
# Take a look at the structure to make sure everything looks right!
str(df_m3)
```

*I was having trouble understanding what the module wanted. I didn't realize you could put the different values back into the original df. Great job!*

## Module 5

> NOTE that to install XLConnect, you will need to download & install Java and install.package("rJava") first (on Macs)

You will also need to install *install.packages("httpuv")* for dropbox section in case you see a bad request popup

> NOTE Shared Dropbox links will take you to a webpage that has the data embedded... to get the raw data you need to change the end of the link from dl=0 to dl=1 or raw=1. That's what the next line of code does:
